% !TEX root = ./main.tex
\section{Plate reader Growth Rate Measurements}
Strains are streaked on LB-Agar plates from frozen stocks to obtain single colonies. Single colonies are picked and grown in 3ml of LB \tr{name brand} overnight in an incubator without shaking. In the morning, cultures are diluted 1:100 into 3ml of minimal media and grown in a shaker until they reach an OD600 of 0.4, which is indicative of exponential phase. Doing the pre-growth in an incubator instead of shaker has shown to have cultures that are grown simultaneously reach exponential phase more consistently \tr{cite paper and possibly show a figure of data?}. When cultures reach exponential phase, they are diluted 1:40 into 300µl of growth media in a well plate. For each well plate, growth media and antibiotic is mixed right before the experiment. To reduce the preparation time, for each antibiotic concentration used in the experiments, a stock of tetracycline (Zymo Research, A1004-25) is prepared in 100\% ethanol such that 50µl of stock can be mixed with 6.5ml of growth media to obtain the target concentration. This ensures less variation in antibiotic concentration across replicates. We used Nunc edge 96 well plates (ThermoFisher Scientific catalog number 167425), which have a moat around the wells which can be filled with water to reduce evaporation during growth. Plates are grown in a Biotek Cytation \tr{Check if either 5 or 7} plate reader at 37C, with measurements every 7 min and linear shaking between measurements (check what speed). Each plate is grown for about 24h, or less if growth has stalled.